Forecasting future oil production in Norway and the UK: a general improved methodology

### 2014-07-18 Publications/Forecasting  Future Oil Production/Source/manuscript.tex
Latest version.

Despite the popularity of Hubbert model, the literature says little about its the out-of-sample performance. There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time. As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test has to be started when the peak is already well visible in the data. For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$. Based on each fit, we computed the predicted remaining recoverable reserves.
This back-test allows us to study how the forecast of Hubbert's model evolves over time. A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time. An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects. This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.
For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000. Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which requires the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks. The peak of each of the two logistic functions matches up with one of the two peaks in the UK production data.
The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time. In 2000, for Norway, the Hubbert model would have predicted 2.08Gb remaining recoverable reserves in 2014, while up to now the predicted amount grew to 2.83Gb (a 36\% error) and the growth did not yet slow down. For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth. These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast. It has to be added that the error in total production since the beginning in the year 2000 relative too now is only 17\% for Norway and 6.5\% for the UK. Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.
The aim is to understand the cause why the Hubbert model underestimates the tail of the production curve and improve upon it. A first argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves. To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price. The forecast for Norway remains fairly stable and then starts to grow with a small delay when the oil price increases. The growth is continuing in the past years, even as the oil price stabilized again. For the UK the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in the oil price. One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to the oil price as new capacities had first to be build. On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price. Any case, it seems difficult to establish a clear link between the oil price and the increase in forecasted reserves.

@@

The existing forecasts of future oil production all use some form of the
Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
or some economical model applied to aggregate production \citep{Greiner2011}.
 However, as just shown in the back-test, the Hubbert model does not provide a
forecast stable in time, usually underestimating the tail in the production curve.
Considering that none of these Hubbert bases methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data. The Hubbert model will certainly always remain the tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.

@@

The first step to predicting the future oil production of a country
is to extrapolate the future production of existing fields and to
estimate the error on this extrapolation. The data of the fields developed
in the past shows a repeating asymmetric pattern. A good example is
the Oseberg field shown in figure \ref{regular}, with a quick ramp
up once the field is being developed, and then a peak or plateau before
the oil field production starts decaying. The decay can take many different shapes
and is governed by a variety of geological and economical factors.
The goal of the fitting procedure is to capture as much the impact of
these different factors as possible.

@@

To be able to forecast the oil production of each individual field,
regularity had to be found in the production's dynamics. Modeling
the whole production profile from the beginning of extraction seems
elusive due to the variety of the forms it can take. Fortunately,
modeling the decay process is sufficient in order to extrapolate future
oil production. A preliminary classification is necessary to achieve
that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
show that, independent of the country, oil fields can be classified
into three main categories:
\begin{itemize}
\item \textbf{Regular fields} - Their decays show some regularity (see figure \ref{regular} ); 
\item \textbf{Irregular fields} - The ones that do not decay in a regular
fashion (see figure \ref{irregular});
\item \textbf{New fields} - The ones that do not decay yet. As such, there
is no easy way to forecast their future oil production based on past
data (see figure \ref{new}).
\end{itemize}
\noindent All the fields have been fitted using an automated algorithm,
but the results have been subsequently checked visually to sort out
the irregular fields which could not be fitted. As of January 2014,
regular fields make up 85\% and 87\% of the number of fields and 94\%
and 71\% of the total produced oil volume in Norway and the U.K. respectively.
As such, being able to model them is crucial. To capture as many different
decay dynamics as possible, the decay part of the oil production rate
$f(t) := \frac{dP}{dt}$ has been fitted by the stretched exponential  
\begin{equation}
f(t)=f_{0}~e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
\label{stretched_exponential}
\end{equation}
The stretched exponential function
has many advantages as it generalises the power law and can therefore
capture a broad variety of distributions as shown by \citet{Laherrere1998}. Moreover,
\citet{Malevergne2005} showed that the power law
function can be obtained as an asymptotic case of the stretched
exponential family, allowing for asymptotically nested statistical tests.  As can be seen
in figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
is a good functional form to fit the decay process of regular fields.

@@

To determine how well the extrapolation based on the stretched exponential
predicts the future production, a complete back-testing has been performed
on each field. A single back-test is made as follows:
\begin{itemize}
\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field is
truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
$T$ is the time counted in months since the production start of the field.
\item The extrapolation of the oil production rate
$f(t) := \frac{dP}{dt}$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{T}\}$.
\item The future production predicted by the extrapolation function $f(t)$ can be
compared to the actual production from the date $T$ in the past
up to the present $T_{f}=N$. The extrapolated total production can be
computed as 
\begin{equation}
P_{e}(T)=\int_{T}^{T_{f}}f(t)dt
\end{equation}
and the relative error is given by
\begin{equation}
e(T)=\frac{P_{e}(T)-\sum_{i=T}^{T_{f}}p_{i}}{P_{e}(T)},
\label{ryjryukoik}
\end{equation}
where both $P_{e}(T)$ and $e(T)$ are functions of the truncation time $T$.
\end{itemize}
Computing this back-test, for every month in the past since the field production
started decaying, yields a plot showing the evolution of the relative
error over time defined by expression (\ref{oesberg-error}). By construction, the relative
error will tend to zero as the truncation time $T$ approaches the
present. Nonetheless, it is a useful indicator for the stability of
the extrapolation. As can been seen for the Oseberg field in figure \ref{oesberg-error}, the relative
error on future production remained fairly stable during the past
decade.

@@

Based on the methodology described in section \ref{sub:Discovery-rate-of},
simulating future oil production was straightforward. For each country,
the existing oil field productions were extrapolated and the future discoveries were
simulated. Figure \ref{no_uk_2013} shows the average of 1000 simulations.
For each country, the non-symmetric shape of the production dynamics, which
contradicts the prediction based on Hubbert's standard approach,
is immediately noticeable.

The results in table \ref{forecast-14} show a striking difference
between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $45\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue for us to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$66\%$ compared to the latter.

Which of the two models is more trustworthy? Clearly, the implications
in adopting one methodology over the other are significant. The only
way to answer this question is to back-test them. In other words:
\textquotedbl{}What would each of the models have predicted, had they
been used in the past?\textquotedbl{} The next section addresses
this question and presents the validation step of our approach.

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the results of these tests. Comparing the forecast of both models,
with the oil production of the subsequent 6 years, shows that the
predictive power of the Monte-Carlo model is significantly better than a simple
extrapolation of the aggregate past production. The Monte-Carlo model 
is found essentially right on target, while the extrapolation of past production
(``fit'') is under-estimating the realised production by 19\% and 16\% respectively
for Norway and the U.K. The following table \ref{backtest-08}
summarizes the difference between the two approaches for the back-testing
period.

@@

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels 
predicted by the extrapolation of the  \textbf{Fit}
of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $3.95\cdot10^{9}$  & $5.72\cdot10^{9}$  & $45\%$\tabularnewline
U.K. & $1.79\cdot10^{9}$  & $2.98\cdot10^{9}$  & $66\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

@@

\begin{table}[H]
\caption{Extrapolation of past oil production (``fit'') and prediction
using the Monte-Carlo model are used on the data set truncated
in 2008. Their forecast for the period 2008-2014 is compared to
the actual realised production.}
\centering %
\begin{tabular}{llll}
 & Actual (barrels) & Fit (barrels) & Model (barrels)\tabularnewline
\midrule Norway & $4.05\cdot10^{9}$  & $3.28\cdot10^{9}$  & $3.95\cdot10^{9}$ \tabularnewline
U.K. & $2.40\cdot10^{9}$  & $2.00\cdot10^{9}$  & $2.28\cdot10^{9}$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

@@

\begin{table}[H]
\caption{Remaining oil reserves forecasted for the period 2014-2030 when using
the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and 
the Monte-Carlo model. The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
\centering %
\begin{tabular}{lllc}
 & Fit (barrels) & Model (barrels) & $\Delta$ \tabularnewline
\midrule Norway 08 & $0.76\cdot10^{9}$  & $5.20\cdot10^{9}$  & $-584\%$\tabularnewline
U.K. 08 & $0.76\cdot10^{9}$  & $2.77\cdot10^{9}$  & $-264\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-08}
\end{table}

@@

\begin{table}[H]
\caption{Oil import (bbl/day) at a constant consumption of 1.5M bbl/day for
the U.K. and 0.22M bbl/day for Norway. The import for the E.U. and
Norway is a lower bound based on the changes in the U.K and Norway. Negative
numbers for Norway represent exports.}
\centering %
\begin{tabular}{lllcc}
 & 2014 & 2020 & 2025 & 2030\tabularnewline
\midrule Norway & $-1.23\cdot10^{6}$  & $-0.88\cdot10^{6}$  & $-0.58\cdot10^{6}$  & $-0.43\cdot10^{6}$ \tabularnewline
U.K. & $0.7\cdot10^{6}$  & $0.9\cdot10^{6}$  & $1.0\cdot10^{6}$  & $1.1\cdot10^{6}$ \tabularnewline
E.U.+Norway & $9.8\cdot10^{6}$  & $10.45\cdot10^{6}$  & $10.85\cdot10^{6}$  & $11.1\cdot10^{6}$ \tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{needs}
\end{table}

>>>>

Despite the popularity of the Hubbert model, the literature says little about its out-of-sample performance. There are many ways to address this question and we choose to study the remaining recoverable reserves predicted by the Hubbert Model when going back in time. As the problem of fitting the logistic curve (equation \ref{logistic}) is under-constrained before the peak of $f(t)$, the back-test had to be started when the peak is already well visible in the data. For each time $T$, a month between the start of the back-test and the date of the latest available date, we fitted the Hubbert model to the data truncated at time $T$. Based on each fit, we computed the predicted remaining recoverable reserves in 2014.\\
This back-test allowed us to study how the forecast of Hubbert's model evolves over time. A good model, which captures well the underlying dynamics of the system, should yield a stable forecast over time. An unstable forecast, where the model constantly has too adapt to new data, can be considered as an indicator that the model is missing some aspects. This is of course assuming that the dynamics of the system are deterministic to higher degree than the error the model is making in its forecast.\\
For the studied cases, namely Norway and the UK, a good starting date was chosen to be the 01/2000. Fitting Norway's production could be made using the simple logistic equation \ref{logistic}, however for the UK the production exhibits two clear peaks, which required the use of a double-logistic function which is the sum of two independent logistic functions with time shifted peaks. The peak of each of the two logistic functions is matched up with one of the two peaks in the UK production data.\\
The results of this back-test can be seen in figure \ref{hubbert-backtest} and clearly show that the forecast is not stable over time. In 2000, for Norway, the Hubbert model would have predicted 2.08Gb remaining recoverable reserves in 2014, while up to now the predicted amount grew to 2.83Gb (a 36\% error) and the growth did not yet slow down. For the UK, the prediction would have been 0.21Gb but grew all the way to 1.43Gb (an error of 581\%), with a recent slow down in growth. These results clearly show that the Hubbert model can be highly unstable and unreliable in its forecast. For completeness, it has to be added that the relative error between the year 2000 and now in total production since the beginning is only 17\% for Norway and 6.5\% for the UK. Nonetheless, the Hubbert model greatly underestimates the fat tail of the production curve.\\
The aim is to understand the causes why the Hubbert model underestimates the tail of the production curve and improve upon it. An argument could be the strong increase in the oil price, which widened the amount of economically recoverable reserves. To discuss this hypothesis the figure \ref{hubbert-backtest} shows as well the oil price. The forecast for Norway remains fairly stable in the beginning and then starts to grow with a small delay when the oil price increases. The growth is continuing in the past years, even as the oil price stabilized again. For the UK the behaviour is different, with increases in the forecast showing very little delay with respect to the increases in the oil price. One could speculate that Norway had little spare capacities and therefore the increase in remaining reserves was always delayed with respect to the oil price as new capacities had first to be build. On the other hand, UK could have had available or upcoming spare capacities which allow to respond rapidly to the increase in the oil price. While it is certain that increases in the oil price will increase the recoverable reservers, it seems difficult to establish a mathematically defined link between the oil price and the increase in forecasted reserves. Consequently, an extension of the Hubbert model integrating the oil price has not been further explored.

@@

The existing forecasts of future oil production all use some form of the
Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
or some economical model applied to aggregate production \citep{Greiner2011}.
 However, as just shown in the back-test, the Hubbert model does not provide a
forecast stable in time, usually underestimating the tail in the production curve.
Considering that none of these Hubbert based methodologies go into the details of studying the underlying dynamics, it is not surprising that some of the dynamics have been missed.
In defense of the existing studies, the main reason for the lack of details is certainly the lack of available data. The Hubbert model will certainly always remain the tool of choice to estimate the total reserves when only the aggregated production of a region (typically a country) is available.

@@

The first step to predicting the future oil production of a country
is to extrapolate the future production of existing fields and to
estimate the error on this extrapolation. The data of the fields developed
in the past shows a repeating asymmetric pattern. A good example is
the Oseberg field shown in figure \ref{regular}, with a quick ramp
up once the field is being developed, and then a peak or plateau before
the oil field production starts decaying. The decay can take many different shapes and is governed by a variety of geological and economical factors.
The goal of the fitting procedure is to capture as much as possible the impact of
these different factors.

@@

To be able to forecast the oil production of each individual field,
regularity had to be found in the production's dynamics. Unfortunately, the whole production profile can take a large variety of shapes with often very asymmetric shapes. Therefore, modeling the whole production profile from the beginning of extraction with the Hubbert model seems elusive. It would require the use of various extensions of the Hubbert model and for each field the most appropriate extension would need to be determined. This is impractical when dealing with hundreds of oil field.\\
Fortunately, modeling the decay process is sufficient in order to extrapolate future oil production. A preliminary classification is necessary to achieve that goal. Figures \ref{regular}, \ref{irregular} and \ref{new} show that, independent of the country, oil fields can be classified into three main categories:
\begin{itemize}
\item \textbf{Regular fields} - Their decays show some regularity (see figure \ref{regular} ); 
\item \textbf{Irregular fields} - The ones that do not decay in a regular
fashion (see figure \ref{irregular});
\item \textbf{New fields} - The ones that do not decay yet. As such, there
is no easy way to forecast their future oil production based on past
data (see figure \ref{new}).
\end{itemize}
\noindent All the fields have been fitted using an automated algorithm,
but the results have been subsequently checked visually to sort out
the irregular fields which could not be fitted. As of January 2014,
regular fields make up 85\% and 87\% of the number of fields and 94\%
and 71\% of the total produced oil volume in Norway and the U.K. respectively.
As such, being able to model them is crucial. To capture as many different
decay dynamics as possible, the decay part of the oil production rate
$f(t) := \frac{dP}{dt}$ has been fitted by the stretched exponential  
\begin{equation}
f(t)=f_{0}~e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
\label{stretched_exponential}
\end{equation}
The stretched exponential function
has many advantages as it generalises the power law and can therefore
capture a broad variety of distributions as shown by \citet{Laherrere1998}. Moreover,
\citet{Malevergne2005} showed that the power law
function can be obtained as an asymptotic case of the stretched
exponential family, allowing for asymptotically nested statistical tests.  As can be seen
in figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
is a good functional form to fit the decay process of regular fields.

@@

For every extrapolation, be it the Hubbert model applied to aggregate production at the country level or the stretched exponential applied to a field's decay, we want to determine how good and stable the predictive power is. To do so, we perform a complete back-testing in every case. This includes a monthly back-test for the extrapolation, using a stretched-exponential, of the decay process of every oil field.\\
A single back-test is made as follows:
\begin{itemize}
\item The production data $\{p_{0},\,\ldots,\, p_{N}\}$ of the field or country is truncated at a certain date in the past $T \in\{0,\,\ldots,\, N\}$, where
$T$ is the time counted in months since the production start of the field.
\item The extrapolation of the oil production rate
$f(t) := \frac{dP}{dt}$ is made based on the truncated data $\{p_{0},\,\ldots,\, p_{T}\}$.
\item The future production predicted by the extrapolation function $f(t)$ can be
compared to the actual production from the date $T$ in the past
up to the present $T_{f}=N$. The extrapolated total production can be
computed as 
\begin{equation}
P_{e}(T)=\int_{T}^{T_{f}}f(t)dt
\end{equation}
and the relative error is given by
\begin{equation}
e(T)=\frac{P_{e}(T)-\sum_{i=T}^{T_{f}}p_{i}}{P_{e}(T)},
\label{ryjryukoik}
\end{equation}
where both $P_{e}(T)$ and $e(T)$ are functions of the truncation time $T$.
\end{itemize}
Computing this back-test, for every month in the past since the production
started decaying, yields a plot showing the evolution of the relative
error over time defined by expression (\ref{oesberg-error}). By construction, the relative
error will tend to zero as the truncation time $T$ approaches the
present. Nonetheless, it is a useful indicator for the stability of
the extrapolation. As can been seen for the Oseberg field in figure \ref{oesberg-error}, the relative
error on future production remained fairly stable during the past
decade.

@@

Based on the methodology described in section \ref{sub:Discovery-rate-of},
simulating future oil production was straightforward. For each country,
the existing oil field productions were extrapolated and the future discoveries were simulated. Figure \ref{no_uk_2013} shows the predicted production resulting from the extrapolation of existing fields and the average of 1000 simulations of future discoveries.
For each country, the non-symmetric shape of the production dynamics, which
contradicts the prediction based on Hubbert's standard approach, is immediately noticeable.

The results in table \ref{forecast-14} show a striking difference
between the Hubbert model, the extrapolation of the fit and the Monte-Carlo model forecast. The preliminary study discussed in the introduction already showed that the Hubbert model forecast was not stable over time and therefore unlikely to be accurate. Indeed the results confirm that the Hubbert forecast is smaller then fitting the decaying process and only half of what the Monte-Carlo model predicts. This is not surprising, because the Hubbert curve is symmetric and therefore the fitting procedure will be strongly biased by the rapid increase in production before the peak, therefore predicting a rapid decrease. Only fitting the decay process using a stretched exponential resolves this problem, because the decay is entirely decoupled from the increase before the peak. Consequently, we expect the simple fit of the decay to perform better.\\
According to the fit (extrapolation of aggregate production), Norway's future oil production would still decay much faster than in the Monte-Carlo case. The remaining reserves estimated with the Monte-Carlo methodology are $45\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue for us to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the fit extrapolation is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$66\%$ compared to the latter.

Which of the models is more trustworthy? Clearly, the implications
in adopting one methodology over the other are significant. The only
way to answer this question is to back-test them. In other words:
\textquotedbl{}What would each of the models have predicted, had they
been used in the past?\textquotedbl{} The next section addresses
this question and presents the validation step of our approach.

@@

For both countries, namely Norway and the U.K., a back-test using
the data truncated in 2008 has been made. Before that date, too many
of the giant fields have not entered their decay phase for a sufficiently
long time to apply the extrapolation algorithm. Figure \ref{model-back-testing}
shows the results of these tests. Comparing the forecast of both models,
with the oil production of the subsequent 6 years, shows that the
predictive power of the Monte-Carlo model is significantly better than the Hubbert model or a simple extrapolation of the aggregate past production. The Monte-Carlo model is found essentially right on target, while the extrapolation of past production (``fit'') is under-estimating the realised production by 19\% and 16\% respectively for Norway and the U.K. The following table \ref{backtest-08} summarizes the difference between the two approaches for the back-testing period.

@@

\begin{table}[H]
\caption{Remaining oil reserves in billion barrels (Gb) until 2030 predicted by the extrapolation of the  \textbf{Fit} of the past country production, and predicted by the Monte-Carlo \textbf{Model}. The relative difference between these two predictions is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
\centering %
\begin{tabular}{llllr}
 & Hubbert (Gb) & Fit (Gb) & Model (Gb)  & $\Delta$ \tabularnewline
\midrule Norway & $2.83$ & $3.95$ & $5.72$ & $45\%$\tabularnewline
U.K. & $1.43$ & $1.79$  & $2.98$ & $66\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

@@

\begin{table}[H]
\caption{Extrapolation of past oil production (``fit'') and prediction
using the Monte-Carlo model are made on the data set truncated
in 2008. Their forecast for the period 2008-2014 is compared to
the actual realised production. All units in billion barrels (Gb).}
\centering %
\begin{tabular}{lllll}
 & Actual (Gb) & Hubbert (Gb) & Fit (Gb) & Model (Gb)\tabularnewline
\midrule Norway & $4.05$  & $3.29$  & $3.28$  & $3.95$ \tabularnewline
U.K. & $2.40$  & $1.63$  & $2.00$  & $2.28$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{backtest-08}
\end{table}

@@

\begin{table}[H]
\caption{Remaining oil reserves in billion barrels (Gb) forecasted for the period 2014-2030 when using the data truncated in 2008, according to the extrapolation of past oil production (``fit'') and the Monte-Carlo model. The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
\centering %
\begin{tabular}{llllc}
 & Hubbert (Gb) & Fit (Gb) & Model (Gb) & $\Delta$ \tabularnewline
\midrule Norway 08 & $2.21$ & $0.76$ & $5.20$  & $-584\%$\tabularnewline
U.K. 08 & $0.73$ & $0.76$ & $2.77$ & $-264\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-08}
\end{table}

@@

\begin{table}[H]
\caption{Oil import in million barrels per day (Mb/day) at a constant consumption of 1.5 Mb/day for the U.K. and 0.22 Mb/day for Norway. The import for the E.U. and Norway is a lower bound based on the changes in the U.K and Norway. Negative numbers for Norway represent exports.}
\centering %
\begin{tabular}{lllcc}
 & 2014 & 2020 & 2025 & 2030\tabularnewline
\midrule Norway (Mb/day) & $-1.23$  & $-0.88$  & $-0.58$  & $-0.43$ \tabularnewline
U.K. (Mb/day) & $0.7$  & $0.9$  & $1.0$  & $1.1$ \tabularnewline
E.U. + Norway (Mb/day) & $9.8$  & $10.45$  & $10.85$  & $11.1$ \tabularnewline
\bottomrule  &  &  &  & \tabularnewline
\end{tabular}\label{needs}
\end{table}
